Recovering from population extinction in the Animal Life Cycle Algorithm (ALCA)

### 2022-09-23 paper.tex
Updated Proposal
Work in progress. Starting with Section 3 Experiments

%PREVIOUSLY / FROM THE FIRST CHAPTER
%We designed our algorithm using observation, analysis, and abstraction in nature to identify what most species in the animal kingdom have in common, where individuals of a population that best adapt to the environment have greater chances of survival, reproduction, and improvement of the species. Imagining as if the task of writing these rules of evolution was in our hands, questioning how we could solve it?
%Like nature, we identified what most species have in common in the life cycle. Our algorithm works on a constantly evolving population, experiencing the stages that all living beings undergo \cite{read1968system}: being born, growing up, reproducing, and dying, where each of these stages will be processes (of our algorithm) that randomly affect the population individuals, emulating the organic way.

@@

Many of the publications with the most influence on our research
~\cite{valdez2021container,garcia2021event,merelo2016performance,merelo2016nodio}
propose a cloud-native optimization architecture that features population-based
algorithms. They present ideas such as those mentioned in Figure~\ref{fig.background},
which planted the seed of thought of our current paper.

\begin{figure}
    \centering
    \includegraphics[width=80mm]{img/fig1_background.pdf}
    \caption{These ideas planted the seed of thought in our algorithm proposal.} \label{fig.background}
    \end{figure}

Our main inspiration for this algorithm was born from the study and observation
of nature, where many successful optimization algorithms have previously found
their own. Our focus was not on a single animal species but all of them, using
general abstraction to identify what those species have in common.

Some questions in our train of thought were: How does nature work? Is evolution
real? How do we enforce natural selection in evolution? What is the role of the
attraction between couples on reproduction? How much of a factor is the
attraction's role in the evolution of the species? What is the role of death in
all of this? How much of an impact does death have on the evolution of a
species? How do we include those ideas into our algorithm?

In our analysis, we consider a broad animal spectrum, from bacteria to humans
and lions, finding our inspiration on the biological Life Cycle of the animal
species, which consists of several stages \cite{read1968system}: birth, growth,
reproduction, and death. As in nature, we intend to execute all these stages in
parallel and asynchronously on a population that evolves constantly. The
combination of those processes is what we call evolution.

One of the main challenges when designing the algorithm was how to capture the
essence of life by reflecting a population evolution?. In our minds, it was
clear this task requires a combination of simultaneous working forces that
influences the population improvement over time. Our strategy was simple:
divide and conquer, looking into the clouds. Enter cloud-computing: previous
cloud-computing algorithms have proven successful results working in
optimization problems \cite{valdez2021container,garcia2021event}.

We turned back to our genetic algorithms knowledge, mixed with our inspiration
and cloud computing, and designed the algorithm with the quest to divide the
processing workload among multiple computers. This strategy would also make it
possible to run our algorithm as a cloud service. Theoretically, one
consequence of the workload distribution is to obtain lower execution times.

This algorithm seeks to imitate the natural life cycle, where new individuals
are born at any moment and mature over time, where they age and suffer
mutations throughout their lives. In reproduction, couples match by mutual
attraction, and they may have offspring. Death can happen to everyone: from a
newborn to an aged adult, where the individual's fitness will impact their
longevity. The general model concept is shown in Figure~\ref{fig.proposal}. The general
flowchart for the Life-Cycle algorithm is described in Figure~\ref{fig.flowchart}.

\begin{figure}
    \centering
    \includegraphics[width=70mm]{img/fig2_proposal.pdf}
    \caption{Algorithm model inspired by the animal species' biological Life-Cycle.} \label{fig.proposal}
    \end{figure}

\begin{figure}
    \includegraphics[width=\textwidth]{img/fig4_flowchart.pdf}
    \caption{Life-Cycle algorithm general flowchart.} \label{fig.flowchart}
% Several terms are introduced here that are not defined in the text, like "champion list" and "attraction" - JJ
    \end{figure}

\subsection{Birth}

This algorithm starts with a randomly generated population, where the processes
will interact with the population independently. This means that at any given
time, any individual can experience any of these processes. Birth is the first
process of our algorithm, and it is responsible for the initial generation of
individuals.

\subsection{Growth}

As in nature, all individuals constantly grow, mature, or age. With increasing
age, individuals may lose strength but also gain more knowledge to solve
problems. We represent this with a possible mutation in each increment of age.
The growth process will take an individual to assess whether it's ready to
mature and undergo changes. % Not learning?  JJ

\subsection{Reproduction}

The attraction of a couple will depend on fitness: the better individual's
fitness, the more attractive it will be, making it easier to find mating
matches. This process will be taking random pairs of individuals to evaluate
their attraction as a couple, to try to breed; when the gestation is
successful, a new pair of individuals will be born (as the offspring). Not all
couples will be compatible, so reproduction will not always be possible, but
the problem arises: how to quantify the attraction between two individuals?

We could have used several strategies to evaluate this attraction. Considering
this algorithm takes a similar focus as the study of bacteria growth in
microbiology, where we can observe and analyze the evolution of a population
over time. As in nature' most species, the bigger a specimen is, the more
attractive it is to its mating candidates. To be consistent with both ideas in
% Er... not really. It depends on health, mainly - JJ
our algorithm, we use the equation of Newton's Law of Universal Gravitation (1)
to calculate the attraction between two individuals, where we visualize the
individual fitness as its mass when using the equation. In previous work,
Newton's Law of Universal Gravitation has shown success in helping to solve
optimization problems \cite{sanchez2014fuzzy,rashedi2009gsa}.

\begin{figure}
    \centering
    \includegraphics[width=50mm]{img/fig3_newtonslaw.pdf}
    \caption{Newton's Law of Universal Gravitation.} \label{fig.newtonslaw}
    \end{figure}

% to be honest, I find this a bit far fetched. How do you define mass? is it related to fitness? - JJ

\subsection{Death}

The death stage represents the challenges and adversities that life presents to
overcome. This process evaluates the individual resistance to survive in the
environment. The better fitness the individual has will increase its chances of
survival. As time progresses, the demands of nature will also increase, pushing
for only the best individuals to survive.

@@

>>>>

@@

Our inspiration for this algorithm was found on the study and observation of
nature, using general abstraction to identify what most animal species have in
common: the life cycle. We present an algorithm inspired by the biological
life-cycle of animal species, which consists of several stages
\cite{read1968system}: birth, growth, reproduction, and death. The combination
of those processes results in evolution. As in nature, we intend to execute all
these stages in parallel and asynchronously on a constantly evolving
population.

\subsection{Algorithm Model}

This algorithm seeks to imitate the natural life cycle, where new individuals
are born at any moment and mature over time, where they age and then suffer
mutations throughout their lives. In reproduction, couples match by mutual
attraction. Death can happen to everyone: from a newborn to an aged adult,
where the individual's fitness will impact their longevity. We display the
general model concept in Figure~\ref{fig.algorithm_model}.

\subsubsection{Birth} Birth is the first process of our algorithm, and it is
responsible for the initial generation of individuals. This algorithm starts
with a randomly generated population.

\subsubsection{Growth} As in nature, all individuals constantly grow, mature,
or age. With increasing age, individuals may lose strength but may also gain
knowledge to solve problems. We represent this with a possible mutation in each
increment of age.

\subsubsection{Reproduction} This process will select random pairs of
individuals to evaluate their attraction as a couple, where this attraction
will depend on fitness. The better individual's fitness, the more attractive it
will be. Not all couples will be compatible, so reproduction will not always be
possible. When the breeding is successful, a new pair of individuals will be
born (offspring).

\subsubsection{Death} The death stage represents the challenges and adversities
that life presents. This process evaluates the individual resistance to survive
in the environment. The better fitness the individual has will increase its
chances of survival. As time progresses, the demands of nature will also
increase, pushing for only the best individuals to survive.

\begin{figure}
    \centering
    \includegraphics[width=80mm]{img/fig_algorithm_model.pdf}
    \caption{Animal Life Cycle Algorithm model.} \label{fig.algorithm_model}
    \end{figure}

\subsection{Problem}

In this algorithm's reproduction process, we find mating couples by mutual
attraction, where sometimes individuals won't procreate offspring because of
the mate's low appeal. As time passes, the environment could kill its
individuals either because they have low fitness or age, leading to extinction,
meaning there are no more individuals in the population left to reproduce. When
this condition presents, if evaluations are available and termination
conditions enable the algorithm to continue the search, it is required to
create a new set of individuals. We illustrate the general flowchart for the
Life-Cycle algorithm in Figure~\ref{fig.flowchart}.

\begin{figure}
    \includegraphics[width=\textwidth]{img/fig_flowchart.pdf}
    \caption{Animal Life-Cycle Algorithm general flowchart.} \label{fig.flowchart}
    \end{figure}

\subsection{Proposed Solution}

In this article, we test various strategies to generate a new population to
replace individuals that have already died (or were discarded) due to their
aptitude, as happens in the life cycle. We previously used to generate a new
random set of solutions, with the cost of losing all evolution knowledge. The
strategy we describe on this paper uses elites to take advantage of the
historical evolution, especially the best individuals, where we look for the
offspring of the elite set to be somewhat close to them. We use the golden
ratio to compute the distance between new individuals in the population, making
an improved distribution.

We propose one promising alternative to restart the population, to solve
extinction caused by nature's pressure in the Life Cycle algorithm: the
projection of the historically best-found candidate solutions (Elite) towards
the best-found candidate solution (Champion). This article proposes some
strategies to carry out this operation, where the presented distribution
strategy obtained best results.

\subsubsection{How does it work?} Based on the historically best candidate
solution and the elite set, we use the Golden Ratio (shown in
Figure~\ref{fig.golden_ratio}) to compute the distance between new individuals
in the population, making an improved distribution. It randomly selects a
solution from the elite set and uses the Golden Ratio to compute multiple
solutions against the best found candidate.

\begin{figure}
    \centering
    \includegraphics[width=25mm]{img/fig_golden_ratio.pdf}
    \caption{Golden Ratio equation.} \label{fig.golden_ratio}
    \end{figure}

This strategy is how we can calculate the projection of the historically
best-found candidate solutions (Elite) towards the best-found candidate
solution (Champion). To illustrate this concept we use
Figure~\ref{fig.elite_projection}, placing the Elite solution on the far right
and the Champion solution on the left corner, where a higher number of new
solutions will be closer to the Champion.

\begin{figure}
    \centering
    \includegraphics[width=0.65\textwidth]{img/fig_elite_projection.pdf}
    \caption{Projection of the historically Elite solutions towards the Champion.} \label{fig.elite_projection}
    \end{figure}

This process generates new solutions that will be the new population set. It
uses the Golden Ratio to progressively aproximate from an Elite solution
towards the Champion. This strategy facilitates to find apparently hidden
solutions, having an exploitation emphasis.
Figure~\ref{fig.elite_projection_swarm} displays the projection of five Elite
solutions (on the corners) towards the Champion (on the center).

\begin{figure}
    \centering
    \includegraphics[width=0.85\textwidth]{img/fig_elite_projection_swarm.pdf}
    \caption{Projection of five Elite solutions towards the Champion, results in a mini-swarm.} \label{fig.elite_projection_swarm}
    \end{figure}

@@

ALTO / HASTA AQUI VAMOS.

%----------------------------------------------------
%----------------------------------------------------
%PREVIOUSLY / FROM THE FIRST CHAPTER
%We designed our algorithm using observation, analysis, and abstraction in nature to identify what most species in the animal kingdom have in common, where individuals of a population that best adapt to the environment have greater chances of survival, reproduction, and improvement of the species. Imagining as if the task of writing these rules of evolution was in our hands, questioning how we could solve it?
%Like nature, we identified what most species have in common in the life cycle. Our algorithm works on a constantly evolving population, experiencing the stages that all living beings undergo \cite{read1968system}: being born, growing up, reproducing, and dying, where each of these stages will be processes (of our algorithm) that randomly affect the population individuals, emulating the organic way.
%Many of the publications with the most influence on our research ~\cite{valdez2021container,garcia2021event,merelo2016performance,merelo2016nodio} propose a cloud-native optimization architecture that features population-based algorithms. They present ideas such as those mentioned in Figure~\ref{fig.background}, which planted the seed of thought of our current paper.
%Some questions in our train of thought were: How does nature work? Is evolution real? How do we enforce natural selection in evolution? What is the role of the attraction between couples on reproduction? How much of a factor is the attraction's role in the evolution of the species? What is the role of death in all of this? How much of an impact does death have on the evolution of a species? How do we include those ideas into our algorithm?
%One of the main challenges when designing the algorithm was how to capture the essence of life by reflecting a population evolution?. In our minds, it was clear this task requires a combination of simultaneous working forces that influences the population improvement over time. Our strategy was simple: divide and conquer, looking into the clouds. Enter cloud-computing: previous cloud-computing algorithms have proven successful results working in optimization problems \cite{valdez2021container,garcia2021event}.
%We turned back to our genetic algorithms knowledge, mixed with our inspiration and cloud computing, and designed the algorithm with the quest to divide the processing workload among multiple computers. This strategy would also make it possible to run our algorithm as a cloud service. Theoretically, one consequence of the workload distribution is to obtain lower execution times.
%----------------------------------------------------
%Birth: This algorithm starts with a randomly generated population, where the processes will interact with the population independently. This means that at any given time, any individual can experience any of these processes. Birth is the first process of our algorithm, and it is responsible for the initial generation of individuals.
%Growth: As in nature, all individuals constantly grow, mature, or age. With increasing age, individuals may lose strength but also gain more knowledge to solve problems. We represent this with a possible mutation in each increment of age. The growth process will take an individual to assess whether it's ready to mature and undergo changes. 
%Reproduction: The attraction of a couple will depend on fitness: the better individual's fitness, the more attractive it will be, making it easier to find mating matches. This process will be taking random pairs of individuals to evaluate their attraction as a couple, to try to breed; when the gestation is successful, a new pair of individuals will be born (as the offspring). Not all couples will be compatible, so reproduction will not always be possible, but the problem arises: how to quantify the attraction between two individuals?
%------------- We could have used several strategies to evaluate this attraction. Considering this algorithm takes a similar focus as the study of bacteria growth in microbiology, where we can observe and analyze the evolution of a population over time. As in nature' most species, the bigger a specimen is, the more attractive it is to its mating candidates. To be consistent with both ideas in our algorithm, we use the equation of Newton's Law of Universal Gravitation (1) to calculate the attraction between two individuals, where we visualize the individual fitness as its mass when using the equation. In previous work, Newton's Law of Universal Gravitation has shown success in helping to solve optimization problems \cite{sanchez2014fuzzy,rashedi2009gsa}.
%Death: The death stage represents the challenges and adversities that life presents to overcome. This process evaluates the individual resistance to survive in the environment. The better fitness the individual has will increase its chances of survival. As time progresses, the demands of nature will also increase, pushing for only the best individuals to survive.
%----------------------------------------------------

\section{Another Section}
\label{section.another_section}
